\begin{appendices}

%
% The first appendix must be "Self-appraisal".
%
\chapter{Self-appraisal}

\section{Critical self-evaluation}
\subsection{Positives}
By far the strongest aspect of my report is the actual solution I produced.  It performs better than any other openly available solution.  
The code itself has received a great deal of attention and utilizes many software engineering principals to ensure the solution is easy to 
read, extendable and performant:  
\begin{itemize}
    \item Environment variables are used to inject configuration into configuration classes  
    \item Python generators were used to avoid loading many GB of data into memory  
    \item Facades were used to abstract underlying implementations for Storage and Camera interfaces  
    \item Proper experiment tracking was used to manage model versions and run experiments in the cloud with A6000 GPUs
    \item Debuggers and profilers were used to benchmark performance and find bugs
\end{itemize}
The results also demonstrated a rigorous and scientific approach to evaluating my solution, improving upon existing solutions and ensuring there was 
no bias when performing these tests.  Yet I was also clear about where my solution had limitations and expanded on what could be done 
to mitigate those.

\subsection{Negatives}
I believe the report could be written better - it is definitely the weakest aspect. 
Despite some effective use of figures and a decent structure, the language is not up to par with a lot of the published works I have read.  
There were a lot of things I felt I wasn't able to convey clearly and at time the language was not consistent, 
however, this is a skill I am glad to have started to learn.
The results section while detailed a comparison against alternative solutions only provided an end-to-end analysis.  
Many more experiments were actually performed during the development of the solution, to measure different aspects and to make data driven design decisions, but 
these were not presented or conveyed to the reader.
The report also didn't really demonstrate my understanding of the underlying architectures and their differences.  However as this is not what I contributed, it 
perhaps isn't necessary.

\section{Personal reflection and lessons learned}
Firstly, this project has been really fun and I have learnt a lot about what it takes to produce even a simple machine learning solution.  
Seeing the end result was even better and I'm genuinely surprised at how well it worked.

One particular area I Learnt a lot about statistics which I never covered at A-Level.

I was also surprised at how effective transfer learning is and has changed my view of its significance in the future.
I trained over 2000 models and was very thankful to my previous self for finding a decent solution at experiment tracking.  This and the data management aspects 
of the project actually peaked my interest the most and has made me very curious as to how these sorts of problems should be solved at massive scale.


\section{Legal, social, ethical and professional issues}

\subsection{Legal issues}
Legal issues could only really enter my project through the data collection and usage aspects.  By using my own data of games I played with no personal information 
I was able to completely avoid any legal issues.  As I am opening my data and solution to be opensource as I found that I really appreciated when others did so legal 
issues may also enter such as licensing.  Any datasets collected with my tool will be the complete responsibility of the person who captured that information and would 
need to be included in the license.

\subsection{Social issues}
The one social issue I can really see with this project is if it is used for cheating.  I also believe it would be trivial to train my solution on digital games such 
as those on chess.com as the data is abundant and would probably work very well.  This could lead to some bad actors using it to run analysis while playing the game 
and so would ruin the game for those who want to play against real humans.

\subsection{Ethical issues}
Personally, there weren't any serious ethical issues except those that I have already mentioned.  The real ethics come into when the tools and techniques I have used 
throughout this project are used for other purposes.  Deep learning is improving at a rapid rate and because of it's nature it's applicability knows really no bounds.
I believe these techniques can be used for immense good and therefore should be avoided, however, the negative impact it could have if not careful could disastrous and 
something I am very passionate about keeping a very close eye on.

\subsection{Professional issues}
If this solution were to be used in a professional setting, such as recording chess games in competitions, and mistakes were made by the model, then I could see some 
pretty bad issues.  I can't image Carlsen being very happy if he wanted to analyse his game but the inference application failed to record to PGN 100\% correctly. 
This was not the aim of the project and more measures and effort would need to be taken to avoid consequences in professional settings.

%
% Any other appendices you wish to use should come after "Self-appraisal". You can have as many appendices as you like.
%
\chapter{External Material}

\section{Software Libraries}
\begin{itemize}
    \item pytorch (https://pytorch.org/)
    \item GuildAI (https://guild.ai/)
    \item python-chess (https://github.com/niklasf/python-chess)
    \item matplotlib (https://matplotlib.org/)
    \item numpy (https://numpy.org/)
    \item opencv (https://opencv.org/)
    \item pandas (https://pandas.pydata.org/)
    \item seaborn (https://seaborn.pydata.org/)
\end{itemize}

\section{PGN File}
Any external PGN files that were not automatically generated from my solution came from PGN Mentor (https://www.pgnmentor.com/)

%
% Other appendices can be added here following the same pattern as above.
%

\chapter{Demos}
\label{demos}

\begin{figure}[h]
    \centering
    \begin{tabular}{|c|c|c|c|}
        \hline
        Game & Total Moves & PGN 100\% Correct & Video  \\
        \hline
        Carlsen/Wesley & 57 & \checkmark & \verb|https://youtu.be/j59y4qAOjcU| \\
        Carlsen/Rapport & 36 & \checkmark & \verb|https://youtu.be/N_vxSqjUWPs|  \\
        Carlsen/Nakamura & 71 & - & \verb|https://youtu.be/3Ji34t7AqWE| \\
        Carlsen/Aranian & 55 & \checkmark & \verb|https://youtu.be/EHoWjNZyeTE|  \\
        Fool's Mate & 4 & \checkmark & \verb|https://youtu.be/Gr4cr5Uptpo| \\
        \hline
    \end{tabular}
    \caption{Games played with the inference application}
    \label{fig:games}
    \end{figure}

\end{appendices}
